Streaming CityJSON datasets

### 2024-04-19 paper/main.tex
Add why rotterdam size increased, I tested with a dataset

First observe that---contrary to intuition---the size of a dataset serialised as a CityJSONSeq file is around 12\% more compact than serialised as a CityJSON file, and in the case of \textbf{3DBAG} it is 36\% compacter.
The main reason for this is that the indices of the vertices are low integers for each feature (because the lowest index in each feature is always ``0''), and do not increase to several millions.
For instance, the dataset \textbf{Helsinki} contains a total of more than 3 millions vertices, but its largest feature contains only but 2202 vertices.
The fact that many indices are used for representing the geometries (and for the textures) means that if several large numbers are used then the size of the file will grow; if they maximum vertex index is around 2000 for each feature, then the file size will be less than the original file.
Only one dataset sees its size slightly increased when serialised to a CityJSONSeq file (\textbf{Rotterdam}).
This is caused, we believe, by the fact that the dataset has textures that are shared by most of the buildings, and the \texttt{"appearance"} property must be repeated for each of the feature (in the CityJSON file it is only stored once and referenced by all buildings).

@@

\begin{itemize}
  \item We have kept most of the structure of CityJSON so that parsers do not need to be adapted much.
  \item CityJSONSeq is an attractive alternative to CityJSON, in most uses it is better to use since it is compacter, processing it uses less memory and is faster.
  \item it opens up many possibilities for processing large 3dcm by using Unix pipes. A few of the tools developed for CityJSON are already reading/writing to stdin/stdout, so they can be chained together.
  \item cjseqview, cjseqval are two of them, but developing others to eg filter, calculate volumes or repair geometries would be trivial
  \item the nice part is that the tools do not need to be written in the same language, one pipeline can have tools built in Python and Rust and C++ and it all works nicely.
  \item extensions: these should be placed in the metadata or collection
\end{itemize}

>>>>

First observe that---contrary to intuition---the size of a dataset serialised as a CityJSONSeq file is around 12\% more compact than serialised as a CityJSON file, and in the case of \textbf{3DBAG} it is 36\% compacter.
The main reason for this is that the indices of the vertices are low integers for each feature (because the lowest index in each feature is always ``0''), and do not increase to several millions.
For instance, the dataset \textbf{Helsinki} contains a total of more than 3 millions vertices, but its largest feature contains only but 2202 vertices.
The fact that many indices are used for representing the geometries (and for the textures) means that if several large numbers are used then the size of the file will grow; if they maximum vertex index is around 2000 for each feature, then the file size will be less than the original file.

Only one dataset sees its size slightly increased when serialised to a CityJSONSeq file (\textbf{Rotterdam}).
This is caused, primarily, by the fact that the dataset has textures that are shared by most of the buildings, and the \texttt{"appearance"} property must be repeated for each of the feature (in the CityJSON file it is only stored once and referenced by all buildings).
If we remove the textures from the CityJSON file it is \qty{2.1}{\mega\byte} and its CityJSONSeq equivalent is \qty{1.3}{\mega\byte} (which means a difference of 62\%)

@@

\begin{itemize}
  \item We have kept most of the structure of CityJSON so that parsers do not need to be adapted much.
  \item CityJSONSeq is an attractive alternative to CityJSON, in most uses it is better to use since it is compacter, processing it uses less memory and is faster.
  \item CityJSONSeq is an attractive alternative to CityJSON; in most uses it is more compact, its processing requires less memory and it is faster.
  \item We have kept most of the structure of CityJSON so that parsers do not need to be adapted much.
  \item it opens up many possibilities for processing large 3dcm by using Unix pipes. A few of the tools developed for CityJSON are already reading/writing to stdin/stdout, so they can be chained together.
  \item cjseqview, cjseqval are two of them, but developing others to eg filter, calculate volumes or repair geometries would be trivial
  \item the nice part is that the tools do not need to be written in the same language, one pipeline can have tools built in Python and Rust and C++ and it all works nicely.
  \item extensions: these should be placed in the metadata or collection
\end{itemize}
